
#Title: Reproducible Research: Peer Assessment 1
##Author: Firdhoush K
##Date: Wednesday, April 15, 2015
##Output: PA1_template.html

##1. Read "Activity Monitoring data" from current working directory and store to a data frame. Transform the date column from 'character' to 'date' type. Load required packages for graphics & statistical operations.

```{r load_data,echo=TRUE}

 data <- read.csv("repdata_data_activity\\activity.csv",stringsAsFactors = FALSE)
 data$date <- as.Date(data$date,"%Y-%m-%d") ##Coercing to date type.
 require(stats)
 require(ggplot2)
 
```

##2. Total number of steps taken per day. Histogram of the total number of steps taken each day.Report the mean and median of the total number of steps taken per day.

```{r total_per_day,echo=TRUE}

 data <- na.omit(data) ## Removing NA's from the data
 rowsum(data$steps,data$date) ## Total number of steps taken per day
 ## Histogram of the total number of steps taken each day
 hist(rowsum(data$steps,data$date),col="red",xlab="Total number of steps taken per day",main="Histogram")
 s<- summary(rowsum(data$steps,data$date)) ## Computing mean and median of the total number of steps taken per day.
 
```
###The `r s[3]` and `r s[4]` for total number of steps taken per day.  The `r s[2]` and `r s[5]` for total number of steps taken per day.

##3. Average daily activity pattern across 5-minute interval.
```{r average_daily_activity_pattern,echo=TRUE}
  
  x<-aggregate(data[,1],list(interval=data$interval),mean) ## Computing Average daily activity pattern across  
                                                           ##  5-minute interval
  ## Plot of the total number of steps taken each day across 5-minute inerval.
  plot(x,type="l",ylab="Frequency(# of steps taken)",main="Average daily activity pattern across 5-minute interval")
    
```

### 5-minute interval number `r x[order(-x[,2]),][1,1]`,contains the maximum number of steps on average across all the days in the dataset.

##4. Inputting missing values.
```{r missing_values,echo=TRUE}
  
 data <- read.csv("repdata_data_activity\\activity.csv",stringsAsFactors = FALSE)
 data$date <- as.Date(data$date,"%Y-%m-%d")
 m<-sum(is.na(data[,1])) ##Finding number of missing values.
 
 ## Filling missing data with mean from the same interval, where means are stored in variable x(Taken from Section 3)
 for (i in  1:17568) {
       if (is.na(data[i,1]) == "TRUE" ) {
        for (j in  1:288){
          if (data[i,3] == x[j,1] ){
          y<- x[j,2]
          }
        } 
        data[i,1] <- y
    }
   }

 
 s<-summary(rowsum(data$steps,data$date)) ## Computing mean and median of the total number of steps taken per day.
 ## Histogram of the total number of steps taken each day after filling NA's with mean.
 hist(rowsum(data$steps,data$date),col="red",xlab="Total number of steps taken per day",main="Histogram") 

 
```

###The total number of missing values in this dataset is `r m`

###The `r s[3]` and `r s[4]` for total number of steps taken per day. This values doesn't change even after filling the NA's. The `r s[2]` and `r s[5]` for total number of steps taken per day.The quartiles have changed as the input values had an impact on those range of data as NA's are more occuring there.

##5. Patterns between weekdays and weekends.Creation of new factor column "wd" indicating weekend or weekday.

```{r weekday_weekends,echo=TRUE}
  ## Creating new factor column "wd" indicating weekend or weekday.
  data$wd <- "weekday" 
  data$wd[weekdays(data$date) == "Saturday" | weekdays(data$date) == "Sunday" ] <- "weekend"
  z<-aggregate(data[,1],list(interval=data$interval,week=data$wd),mean)## Computing Average daily activity pattern 
                                                                       ## cross  5-minute interval 
          
  ggplot(z,aes(interval,x)) + geom_line() + facet_grid(.~week) + ylab("Number of steps")
    
  
```
